\vspace{-2.5em} Measuring the angular dipole spectrum with arbitrary\\ polarized illumination microscopes\vspace{-1em}

### 2019-02-21 notes/2018-12-10-angular-svd/report/report.tex
Latest angular notes with summary.

The detection kernel is given by
\begin{align}
  h^{\text{det}}(\sd,\so) = \ints{2}d\mh{s}\,\Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)(\mh{s}\cdot\sd)^{-1/2}[1 - (\mh{s}\cdot\so)^2], \label{eq:detkern}
\end{align}
where the $[1 - (\mh{s}\cdot\so)^2]$ term models the familiar $\sin^2\vartheta$
dependence of dipole radiation where $\vartheta$ is the angle between $\sd$ and
$\so$, the $(\mh{s}\cdot\sd)^{-1/2}$ term models the apodization of an aplanatic
objective, and the integral over
$\Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)$ models the finite aperture
of the objective with $\text{NA} = n_o\sin\alpha$. In Appendix
\ref{sec:spherical} we evaluate this integral in closed form to find that
\begin{align}
  h^{\text{det}}(\sd\cdot\so) \propto (1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right]P_2(\sd\cdot\so).
\end{align}
The detection kernel is rotationally symmetric, so it can be written as a
function of $\sd\cdot\so$. For very low NA the detection kernel reduces to
$[1 - (\mh{s}\cdot\so)^2]$ as we would expect.

@@

Our remaining task is to calculate the dipole angular transfer function. We can
start by expanding the kernel into excitation and detection parts
\begin{align}
  H_\ell^m(\sd, \pp) = \ints{2}d\so\, h^{\text{exc}}(\pp\cdot\so)h^{\text{det}}(\sd\cdot\so)Y_\ell^{m*}(\so). 
\end{align}
We can expand both the excitation and detection kernels into a sum of spherical harmonics (a Laplace series)
\begin{align}
    H_\ell^m(\sd, \pp) = \ints{2}d\so\, \left[\sum_{\ell' m'}H_{\ell'}^{m',\text{exc}}(\pp\cdot\so)Y_{\ell'}^{m'}(\so)\right]\left[\sum_{\ell'' m''}H_{\ell''}^{m'',\text{det}}(\sd\cdot\so)Y_{\ell''}^{m''}(\so)\right]Y_\ell^{m*}(\so),\label{eq:triple}
\end{align}
where $H_{\ell'}^{m',\text{exc}}(\pp\cdot\so)$ and $H_{\ell''}^{m'',\text{det}}(\sd\cdot\so)$ are the spherical harmonic transforms of the excitation and detection kernels
\begin{align}
  H_{\ell'}^{m',\text{exc}}(\pp) &= \ints{2}d\so\, h^{\text{exc}}(\pp\cdot\so)Y_{\ell'}^{m'*}(\so),\label{eq:sh1}\\
  H_{\ell''}^{m'',\text{det}}(\sd) &= \ints{2}d\so\, h^{\text{det}}(\sd\cdot\so)Y_{\ell''}^{m''*}(\so)\label{eq:sh2}.
\end{align}
In Appendix B we evaluate these spherical Fourier transforms as
\begin{align}
    H_{\ell'}^{m',\text{exc}}(\pp) &\propto Y_{\ell'}^{m'*}(\pp)\left[\delta_{0\ell'} + \frac{2}{5}\delta_{2,\ell'}\right],\label{eq:dipexc}\\
    H_{\ell''}^{m'',\text{det}}(\sd) &\propto Y_{\ell''}^{m''*}(\sd)\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \left(\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right)\delta_{2,\ell''}\right].\label{eq:dipdet}
\end{align}

@@

\section{Specific designs (in progress)}
one-view epi,
one-view lightsheet,
two-view lightsheet,
two-view lightsheet two-tilt one-side,
two-view lightsheet two-tilt two-side,
two-view lightsheet three-tilt one-side two-tilt one-side.

@@

We can use this result to evaluate the detection kernel
\begin{align}
  h^{\text{det}}(\sd,\so) = \ints{2}d\mh{s}\,\Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)(\mh{s}\cdot\sd)^{-1/2}[1 - (\mh{s}\cdot\so)^2]. \label{eq:kern}
\end{align}
Comparing Eq. \eqref{eq:kern} to Eq. \eqref{eq:integral} means that
\begin{align}
  f(\mh{s}\cdot\sd) &= 1 - (\mh{s}\cdot\so)^2,  \\
  h(\mh{s}\cdot\sd) &= \Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)(\mh{s}\cdot\sd)^{-1/2}.
\end{align}
Evaluating the Fourier-Legendre transform of these functions gives
\begin{align}
  F_\ell &= \frac{2\ell + 1}{2}\int_{-1}^1dx\,(1 - x^{2})P_\ell(x) = \frac{2}{3}(\delta_{\ell,0} + \delta_{\ell,2}),\label{eq:fl}\\
  H_\ell &= \frac{2\ell + 1}{2}\int_{\cos\alpha}^1dx\,x^{-1/2}P_\ell(x) = (1 - \sqrt{\cos\alpha})\delta_{\ell,0} + \left[\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right]\delta_{\ell,2}+ \cdots,\label{eq:hl}
\end{align}
Notice that we only need to evaluate the first two terms of $H_\ell$ since
only two terms of $F_\ell$ are non-zero. As we've noted before, the physical
dipole emission process is the band limiting step here---not the detection
process.

Plugging Eqs. \eqref{eq:fl} and \eqref{eq:hl} into Eq. \eqref{eq:five} gives the
closed form result for the detection kernel
\begin{align}
  h^{\text{det}}(\sd\cdot\so) \propto (1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right]P_2(\sd\cdot\so).
\end{align}

@@

We can use this result to evaluate the excitation and detection angular transfer
functions
\begin{align}
  H_{\ell'}^{m',\text{exc}}(\pp\cdot\so) &= \ints{2}d\so\, h^{\text{exc}}(\pp\cdot\so)Y_{\ell'}^{m'*}(\so),\\
  H_{\ell''}^{m'',\text{det}}(\sd\cdot\so) &= \ints{2}d\so\, h^{\text{det}}(\sd\cdot\so)Y_{\ell''}^{m''*}(\so).
\end{align}
where
\begin{align}
  h^{\text{exc}}(\pp\cdot\so) &\propto (\pp\cdot\so)^2, \\
  h^{\text{det}}(\sd\cdot\so) &\propto (1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right]P_2(\sd\cdot\so).
\end{align}
Starting with the excitation kernel
\begin{align}
  H_{\ell'}^{m',\text{exc}}(\pp) &= \frac{4\pi}{2\ell + 1}Y_{\ell'}^{m'*}(\pp)\left[\frac{2\ell' + 1}{2}\int_{-1}^1dx\, x^2P_{\ell'}(x)\right],\\
  H_{\ell'}^{m',\text{exc}}(\pp) &=\frac{4\pi}{2\ell + 1}Y_{\ell'}^{m'*}(\pp)\left[\frac{1}{3}\delta_{0\ell'} + \frac{2}{3}\delta_{2,\ell'}\right],\\
  H_{\ell'}^{m',\text{exc}}(\pp) &\propto Y_{\ell'}^{m'*}(\pp)\left[\delta_{0\ell'} + \frac{2}{5}\delta_{2,\ell'}\right].
\end{align}
Notice that the polarizer orientation $\pp$ only affects the $m$ components of
the angular transfer function while the relative contribution to the two $\ell$ bands is fixed.

Finally, the detection kernel is
\begin{align}
  H_{\ell''}^{m'',\text{det}}(\sd) &= \frac{4\pi}{2\ell'' + 1}Y_{\ell''}^{m''*}(\sd)\left[\frac{2\ell'' + 1}{2}\int_{-1}^1dx\, h^{\text{det}}(x)P_{\ell''}(x)\right],\\
  H_{\ell''}^{m'',\text{det}}(\sd) &= \frac{4\pi}{2\ell'' + 1}Y_{\ell''}^{m''*}(\sd)\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \frac{1}{5}\left(\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right)\delta_{2,\ell''}\right],\\
  H_{\ell''}^{m'',\text{det}}(\sd) &\propto Y_{\ell''}^{m''*}(\sd)\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \left(\frac{1}{4}\sqrt{\cos\alpha}(7-3\cos(2\alpha)) - 1\right)\delta_{2,\ell''}\right].
\end{align}
Notice that the detection axis $\sd$ only affects the $m$ components of the
detection axis and the detection NA affects the relative contribution to the two $\ell$ bands.

>>>>

The detection kernel is given by
\begin{align}
  h^{\text{det}}(\sd,\so) = \ints{2}d\mh{s}\,\Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)(\mh{s}\cdot\sd)^{-1/2}[1 - (\mh{s}\cdot\so)^2], \label{eq:detkern}
\end{align}
where the $[1 - (\mh{s}\cdot\so)^2]$ term models the familiar $\sin^2\vartheta$
dependence of dipole radiation where $\vartheta$ is the angle between $\sd$ and
$\so$, the $(\mh{s}\cdot\sd)^{-1/2}$ term models the apodization of an aplanatic
objective, and the integral over
$\Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)$ models the finite aperture
of the objective with $\text{NA} = n_o\sin\alpha$. In Appendix
\ref{sec:spherical} we evaluate this integral in closed form to find that
\begin{align}
  h^{\text{det}}(\sd\cdot\so) \propto (1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right]P_2(\sd\cdot\so).
\end{align}
The detection kernel is rotationally symmetric, so it can be written as a
function of $\sd\cdot\so$. For very low NA the detection kernel reduces to
$[1 - (\mh{s}\cdot\so)^2]$ as we would expect.

@@

Our remaining task is to calculate the dipole angular transfer function. We can
start by expanding the kernel into excitation and detection parts
\begin{align}
  H_\ell^m(\sd, \pp) = \ints{2}d\so\, h^{\text{exc}}(\pp\cdot\so)h^{\text{det}}(\sd\cdot\so)Y_\ell^{m*}(\so). 
\end{align}
We can expand both the excitation and detection kernels into a sum of spherical harmonics (a Laplace series)
\begin{align}
    H_\ell^m(\sd, \pp) = \ints{2}d\so\, \left[\sum_{\ell' m'}H_{\ell'}^{m',\text{exc}}(\pp\cdot\so)Y_{\ell'}^{m'}(\so)\right]\left[\sum_{\ell'' m''}H_{\ell''}^{m'',\text{det}}(\sd\cdot\so)Y_{\ell''}^{m''}(\so)\right]Y_\ell^{m*}(\so),\label{eq:triple}
\end{align}
where $H_{\ell'}^{m',\text{exc}}(\pp\cdot\so)$ and $H_{\ell''}^{m'',\text{det}}(\sd\cdot\so)$ are the spherical harmonic transforms of the excitation and detection kernels
\begin{align}
  H_{\ell'}^{m',\text{exc}}(\pp) &= \ints{2}d\so\, h^{\text{exc}}(\pp\cdot\so)Y_{\ell'}^{m'*}(\so),\label{eq:sh1}\\
  H_{\ell''}^{m'',\text{det}}(\sd) &= \ints{2}d\so\, h^{\text{det}}(\sd\cdot\so)Y_{\ell''}^{m''*}(\so)\label{eq:sh2}.
\end{align}
In Appendix B we evaluate these spherical Fourier transforms as
\begin{align}
    H_{\ell'}^{m',\text{exc}}(\pp) &\propto Y_{\ell'}^{m'*}(\pp)\left[\delta_{0\ell'} + \frac{2}{5}\delta_{2,\ell'}\right],\label{eq:dipexc}\\
    H_{\ell''}^{m'',\text{det}}(\sd) &\propto Y_{\ell''}^{m''*}(\sd)\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \left(\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right)\delta_{2,\ell''}\right].\label{eq:dipdet}
\end{align}

@@

% \section{Specific designs (in progress)}
% one-view epi,
% one-view lightsheet,
% two-view lightsheet,
% two-view lightsheet two-tilt one-side,
% two-view lightsheet two-tilt two-side,
% two-view lightsheet three-tilt one-side two-tilt one-side.

@@

We can use this result to evaluate the detection kernel
\begin{align}
  h^{\text{det}}(\sd,\so) = \ints{2}d\mh{s}\,\Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)(\mh{s}\cdot\sd)^{-1/2}[1 - (\mh{s}\cdot\so)^2]. \label{eq:kern}
\end{align}
Comparing Eq. \eqref{eq:kern} to Eq. \eqref{eq:integral} means that
\begin{align}
  f(\mh{s}\cdot\sd) &= 1 - (\mh{s}\cdot\so)^2,  \\
  h(\mh{s}\cdot\sd) &= \Pi\left(\frac{\mh{s}\cdot\sd}{\cos\alpha}\right)(\mh{s}\cdot\sd)^{-1/2}.
\end{align}
Evaluating the Fourier-Legendre transform of these functions gives
\begin{align}
  F_\ell &= \frac{2\ell + 1}{2}\int_{-1}^1dx\,(1 - x^{2})P_\ell(x) = \frac{2}{3}(\delta_{\ell,0} + \delta_{\ell,2}),\label{eq:fl}\\
  H_\ell &= \frac{2\ell + 1}{2}\int_{\cos\alpha}^1dx\,x^{-1/2}P_\ell(x) = (1 - \sqrt{\cos\alpha})\delta_{\ell,0} + \left[\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right]\delta_{\ell,2}+ \cdots,\label{eq:hl}
\end{align}
Notice that we only need to evaluate the first two terms of $H_\ell$ since
only two terms of $F_\ell$ are non-zero. As we've noted before, the physical
dipole emission process is the band limiting step here---not the detection
process.

Plugging Eqs. \eqref{eq:fl} and \eqref{eq:hl} into Eq. \eqref{eq:five} gives the
closed form result for the detection kernel
\begin{align}
  h^{\text{det}}(\sd\cdot\so) \propto (1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right]P_2(\sd\cdot\so).
\end{align}

@@

We can use this result to evaluate the excitation and detection angular transfer
functions
\begin{align}
  H_{\ell'}^{m',\text{exc}}(\pp\cdot\so) &= \ints{2}d\so\, h^{\text{exc}}(\pp\cdot\so)Y_{\ell'}^{m'*}(\so),\\
  H_{\ell''}^{m'',\text{det}}(\sd\cdot\so) &= \ints{2}d\so\, h^{\text{det}}(\sd\cdot\so)Y_{\ell''}^{m''*}(\so).
\end{align}
where
\begin{align}
  h^{\text{exc}}(\pp\cdot\so) &\propto (\pp\cdot\so)^2, \\
  h^{\text{det}}(\sd\cdot\so) &\propto (1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right]P_2(\sd\cdot\so).
\end{align}
Starting with the excitation kernel
\begin{align}
  H_{\ell'}^{m',\text{exc}}(\pp) &= \frac{4\pi}{2\ell + 1}Y_{\ell'}^{m'*}(\pp)\left[\frac{2\ell' + 1}{2}\int_{-1}^1dx\, x^2P_{\ell'}(x)\right],\\
  H_{\ell'}^{m',\text{exc}}(\pp) &=\frac{4\pi}{2\ell + 1}Y_{\ell'}^{m'*}(\pp)\left[\frac{1}{3}\delta_{0\ell'} + \frac{2}{3}\delta_{2,\ell'}\right],\\
  H_{\ell'}^{m',\text{exc}}(\pp) &\propto Y_{\ell'}^{m'*}(\pp)\left[\delta_{0\ell'} + \frac{2}{5}\delta_{2,\ell'}\right].
\end{align}
Notice that the polarizer orientation $\pp$ only affects the $m$ components of
the angular transfer function while the relative contribution to the two $\ell$ bands is fixed.

Finally, the detection kernel is
\begin{align}
  H_{\ell''}^{m'',\text{det}}(\sd) &= \frac{4\pi}{2\ell'' + 1}Y_{\ell''}^{m''*}(\sd)\left[\frac{2\ell'' + 1}{2}\int_{-1}^1dx\, h^{\text{det}}(x)P_{\ell''}(x)\right],\\
  H_{\ell''}^{m'',\text{det}}(\sd) &= \frac{4\pi}{2\ell'' + 1}Y_{\ell''}^{m''*}(\sd)\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \frac{1}{5}\left(\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right)\delta_{2,\ell''}\right],\\
  H_{\ell''}^{m'',\text{det}}(\sd) &\propto Y_{\ell''}^{m''*}(\sd)\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \left(\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right)\delta_{2,\ell''}\right].
\end{align}
Notice that the detection axis $\sd$ only affects the $m$ components of the
detection axis and the detection NA affects the relative contribution to the two $\ell$ bands.
